---
title: "opendata_NRBP"
output: html_document
date: '2024-12-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("terra")
library(terra)
#a=rast("Downloads/210_Sentinel2_to_XarrayDS_to_Rectangles_1000m/092_Features_100px_NC_Store/Feature_01523gc_0087lc_20170412T093031_Fire_1.10m_100px_nc")

#install.packages("ncdf4")
library(ncdf4)
library(ggplot2)
library(strucchange)
 library(RColorBrewer) 
library(grid)
```
 
```{r pressure, echo=FALSE}
# direct reading from terra mixes spectral bands and time. 

setwd("~/Documents/work/sentinel2dimexp/")
#a1_fire=nc_open("~/Downloads/210_Sentinel2_to_XarrayDS_to_Rectangles_1000m/092_Features_100px_NC_Store/Feature_01523gc_0087lc_20170412T093031_Fire_1.10m_100px_nc")
 
a1=nc_open("210_Sentinel2_to_XarrayDS_to_Rectangles_1000m/092_Features_100px_NC_Store/Feature_01935gc_0096lc_20171118T093251_Tillage_3.10m_100px_nc")
SCL <- ncvar_get(a1, "SCL") #read a band

B_nir = ncvar_get(a1, "B08") #read a band
B_red = ncvar_get(a1, "B04")  
B_blue = ncvar_get(a1, "B02")  
B_green =ncvar_get(a1, "B03")

NDVI=(B_nir-B_red)/(B_nir+B_red)
NRBP = (B_red-B_blue)/(B_red+B_blue) #NRBP=Red+Blue / Red−Blue​

x <- ncvar_get(a1, "x")
y <- ncvar_get(a1, "y")
timesenti <- ncvar_get(a1, "time")
#r=rast(NDVI, extent=ext(min(x), max(x), min(y), max(y)), crs="+init=epsg:32632")
r=rast(NRBP, extent=ext(min(x), max(x), min(y), max(y)), crs="+init=epsg:32632")
 
#names(r)=timesenti
timesenti_readable=as.POSIXlt(timesenti,origin="2017-01-02 09:34:02")
names(r)=timesenti_readable
```

Visualise the false color images:
Red Areas → Healthy vegetation (forests, croplands, grasslands).
Dark Blue/Black Areas → Water bodies (rivers, lakes, wetlands).
Gray/White/Blue Areas → Urban infrastructure or bare ground.
Light Cyan/White → Clouds, snow, or high-altitude regions.
Brown/Orange Areas → Sparse or stressed vegetation.


```{r eval=T}
rgreen=rast(B_green, extent=ext(min(x), max(x), min(y), max(y)), crs="+init=epsg:32632")

rnir=rast(B_nir, extent=ext(min(x), max(x), min(y), max(y)), crs="+init=epsg:32632")

rred=rast(B_red, extent=ext(min(x), max(x), min(y), max(y)), crs="+init=epsg:32632")

 
rcomp1 = c(rnir,rred,rgreen)
every = 5
for (nn in seq(1,338, by = every)){
plotRGB(rcomp1,
        r = nn,  # NIR in the Red channel
        g = nn+338,  # Red in the Green channel
        b = nn+338*2,  # Green in the Blue channel
        scale = 65535,  # Adjust based on your data range (e.g., 0–255 or 0–65535)
        stretch = "lin",  # Linear stretch for better visualization
        main = as.character(timesenti_readable[nn]))
  title(main =  paste0(as.character(timesenti_readable[nn])," SCL:",as.character(SCL[1,1,nn])))
}

```


```{r plot, echo=FALSE}
every=4 # plot every e.g. 4 images
#pdf(file=paste0("tillageb_ndvi_every",every,".pdf"))
#defalut raster plot 16 each graph
plot(subset(r, seq(1+every*0,nlyr(r),by=every)))
plot(subset(r, seq(1+every*16,nlyr(r),by=every)))
plot(subset(r, seq(1+every*32,nlyr(r),by=every)))
plot(subset(r, seq(1+every*64,nlyr(r),by=every)))
#dev.off() 
```

sample and plot time series
```{r sample}
samplepoints <- spatSample(r, size = 100, method = "random", na.rm = FALSE, xy = TRUE, cells=T)
plot(subset(r,1), main = "Raster with Sampled Points", legend = TRUE)
points(samplepoints[, c("x", "y")], col = "red", pch = 20, cex = 1)
legend("top", legend = "Sampled Points", pch = 20, col = "red",cex=0.8)
```


```{r plotsampletimeseries}

# Print the sampled points
get_sample_ts = function(coor_id, sample_points=samplepoints){
  coords <- sample_points[coor_id , c("x", "y")]
  # Extract the time series (values across all layers) at this location
time_series <- terra::extract(r, coords)
as.numeric(time_series)
  time_steps <-  timesenti_readable
  
  # Combine extracted time series with time steps
  time_series_df <- data.frame(Date = time_steps, Value = as.numeric(time_series[1, -1]))
#  plot(time_series_df, typ="l")
 

# Plot the time series
  return(time_series_df)

}
plot_sample_ts = function(coor_id, sample_points=samplepoints){
   time_series_df =  get_sample_ts(coor_id, sample_points=samplepoints)
gg = ggplot(time_series_df, aes(x = Date, y = Value)) +
  geom_line(color = "blue") +           # Line plot
  geom_point(color = "red") +           # Add points
  labs(
    title = "NDVI Time Series Plot",
    x = "Date",
    y = "NDVI Value"
  ) +
  theme_minimal()  # Clean theme
 #grid::grid.draw(ggplotGrob(gg)) 
}
```

```{r}
random_integers <- sample(1:100, size = 20, replace = FALSE)

 lapply(random_integers, plot_sample_ts,samplepoints)
```
 Quick test using efp.
```{r}
efptest=function(id){
tsdf = get_sample_ts(id)
ts_zoo = zoo(tsdf)
zoo_data <- zoo(tsdf$Value, order.by = tsdf$Date)
 
bp <- efp(zoo_data~1)
plot(bp)
## calculate corresponding test statistic
test = sctest(bp)
return(test$p.value)
}
#if(test$p.value<0.05){
#  id1 = cbind(id1, id)
 # print(paste0(id,":",test$p.value))
#  }
#return(id1)
#}
 
```
```{r}
a = lapply(1:100, efptest)
a1 = as.numeric(a)
which(a1<0.05)
a1[a1 <0.05]
```


